### 2024-01-07 documentation/thesis/thesis_document.tex
more thesis - first external review

\hypertarget{shimming-procedure}{%
\subsection{Shimming procedure}\label{shimming-procedure}}

@@

@@

\ref{state-of-the-art}. \textbf{State of the art} describes the current
state of technology and forms the basis for further development.
Existing measurement methods and findings are shown here in order to
define the context for the current project.

\ref{usecases}. \textbf{Usecases} are the application cases in which the
project is to be used. They illustrate the practical scenarios and
define how the product can be used in the real world. This is crucial
for understanding the needs of the target group and designing the end
result accordingly.

\ref{unified-sensor}. \textbf{Unified Sensor} refers to the integration
of different sensors into a standardised solution. This enables simple
data acquisition and serves as a basic hardware system on which the
subsequent data processing library can be applied.

\ref{software-readout-framework}. \textbf{Software readout framework}
chapter describes the implementation of the data readout framework. This
includes an explanation of the various modules and specific application
examples.

\ref{usability-improvements}. \textbf{Usability improvements} refers to
measures to improve user-friendliness. This involves optimising
interfaces, interactions and processes to ensure intuitive and efficient
use of the product. This includes, code documentation or the
distribution of source code to users.

\ref{evaluation}. \textbf{Evaluation} comprises the systematic review
and assessment of the overall system. This includes the demonstration of
the implemented capabilities of the overall system against the
previously defined use cases.

\hypertarget{state-of-the-art}{%
\chapter{State of the art}\label{state-of-the-art}}

\hypertarget{opensource-projects}{%
\section{Opensource projects}\label{opensource-projects}}

\hypertarget{conceptual-design}{%
\section{Conceptual design}\label{conceptual-design}}

@@

@@

The following section defines some use cases that the future project
should be able to cover. They illustrate practical situations and help
to understand the functionality and added value of the implemented
project for the user. These were defined in the course of project
planning and provide an overview of how the user interacts with the
project and what functionalities can be expected. In the later
evaluation\ref{evaluation} process, the defined usecases are also used
as a reference to demonstrate the implemented capabilities of the
solution.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Ready to use hardware sensor designs}
\item
  \textbf{Taking automatic measurements from sensors}
\item
  \textbf{Ready to use data analysis functions}
\item
  \textbf{Open storage formats for data export}
\item
  \textbf{Easy software configuration}
\item
  \textbf{User expandable data processing pipelines}
\end{enumerate}

@@

The main aim of this project is to develop a low-cost Hall sensor
interface that is also universally expandable. The focus is on mapping
different sensors and being compatible with different magnet types and
shapes. This ensures broad applicability in different scenarios. Another
goal is reproducibility to ensure consistent results. Easy communication
with standard PC hardware maximises the user-friendliness of the
interface. The flexibility to support different sensors and magnets
makes the system versatile and opens up possibilities for different
applications. A low-cost Hall sensor interface will therefore not only
be economically attractive, but also facilitate the integration of Hall
sensors in different contexts. In addition, the low-cost Hall sensor
interface will serve as a development platform for the data evaluation
\gls{mrp} library and provide real measurement data from magnets. In
addition, the interface firmware creates a basis for the development of
a data protocol for exchanging measured values. This makes it easy to
integrate your own measuring devices into the \gls{mrp} ecosystem at a
later date. This is only possible with a minimal functional hardware and
firmware setup and was developed for this purpose first.

\hypertarget{sensor-selection}{%
\section{Sensor selection}\label{sensor-selection}}

@@

Linear analogue Hall sensors were deliberately omitted, although they
are suitable for more precise measurements and extended measuring
ranges. They were excluded due to the more complex circuit layouts
required and more complex power management. In the context of the
desired goal of developing a cost-efficient and universally expandable
Hall sensor interface, the decision in favour of digital sensors seems
appropriate.

\begin{longtable}[]{@{}lllll@{}}
\caption{Implemented digital halleffect sensors
\label{Implemented_digital_halleffect_sensors.csv}}\tabularnewline
\toprule
& TLV493D-A1B6 & HMC5883L & MMC5603NJ & AS5510\tabularnewline
\midrule
\endfirsthead
\toprule
& TLV493D-A1B6 & HMC5883L & MMC5603NJ & AS5510\tabularnewline
\midrule
\endhead
Readout-Axis & 3D & 3D & 3D & 1D\tabularnewline
Temperature-Sensor & yes & no & yes & no\tabularnewline
Resolution {[}uT{]} & 98 & 0.2 & 0.007 & 48\tabularnewline
Range {[}mT{]} & ±130.0 & ±0.8 & ±3 & ±50\tabularnewline
Interface & \gls{i2c} & \gls{i2c} & \gls{i2c} & \gls{i2c}\tabularnewline
\bottomrule
\end{longtable}

Focussing on digital interfaces not only facilitates implementation, but
also contributes to overall cost efficiency. At the same time, the
integration of temperature sensors enables precise measurements under
varying environmental conditions. This strategic choice forms the basis
for a flexible, universally applicable Hall sensor interface that can be
seamlessly integrated into various existing systems.

The table\ref{Implemented_digital_halleffect_sensors.csv} shows a
selection of sensors for which hardware and software support has been
implemented. The resolution of the selected sensors covers the expected
range of values required by the various magnets to be tested.

@@

The mechanical structure of a sensor is kept very simple for the
following tests. The focus was on providing a stable foundation for the
sensor \gls{ic} and an exchangeable holder for different magnets.

The following figure\ref{1D_sensor_mechanical_3D_printed_structure.png},
shows a sectional view of the \gls{cad} drawing of the
\passthrough{\lstinline!1D: Single sensor!}\ref{d-single-sensor}.

All parts were produced using the 3D printing process. The sensor
circuit board was glued underneath the magnet holder. This is
interchangeable, so different distances between sensor and magnet can be
realised.

The exchangeable magnetic holder (shown in green) can be adapted to
different magnets. It can be produced quickly due to the small amount of
material used. The two recesses lock the holder with the inserted magnet
over the sensor. Due to predetermined tolerances, the magnet can be
inserted into the holder with repeat accuracy and without play. This is
important if several magnets have to be measured, where the positioning
over the sensor must always be the same.

@@

As with the mechanical design, the electronics should be kept as simple
as possible so that it is possible to use the microcontroller and a
magnetic field sensor.

The focus here was on utilising existing microcontroller development and
evaluation boards, which already integrate all the components required
for smooth operation. This not only enabled a time-saving
implementation, but also ensured a cost-efficient realisation.

A decisive step in my work was the in-house PCB
design\ref{1D_sensor_schematic_and_circuit_board.png}, which leads out
all the important connections. The placement and alignment of components
optimised the wiring, which in turn improved the reliability of the
sensor tests. Particular attention was paid to the implementation of a
\passthrough{\lstinline!SYNC!}-\gls{gpio}, which enables subsequent
multi-sensor synchronisation. This functionality opens up the
possibility of synchronising data from different sensors to achieve
precise and coherent measurement results. Overall, this integrated
approach represents an effective solution for the flexible evaluation of
sensors and helps to optimise the development process.

@@

Microcontroller firmware is software that is executed on a
microcontroller in embedded systems. It controls the hardware and
enables the execution of predefined functions. The firmware is used to
process input data, control output devices and fulfil specific tasks
according to the program code. It handles communication with sensors,
actuators and other peripheral devices, processing data and making
decisions. Firmware is critical to the functioning of devices.

@@

The firmware is responsible for detecting the possible connected
sensors\ref{Implemented_digital_halleffect_sensors.csv} and reading them
out. This measured data can be forwarded to a host \gls{pc} via a user
interface and can then be further processed there.

An important component is that as many common sensors as possible can be
easily connected without having to adapt the firmware. This modularity
was implemented using \passthrough{\lstinline!abstract!} class design.
These are initiated according to the sensors found at startup. If new
hardware is to be integrated, only the required
functions\ref{lst:CustomSensorClass} need to be implemented.

@@

The flow
chart\ref{Unified_sensor_firmware_simplified_program_strucutre.png}
shows the start process and the subsequent main loop for processing the
user commands and sensor results. When the microcontroller is started,
the software checks whether known sensors are connected to \gls{i2c} or
\gls{uart} interfaces. If any are found (using a dedicated \gls{lut}
with sensor address information), the appropriate class instances are
created and these can later be used to read out measurement results.

Next, the subsystem for multi-sensor
synchronisation\ref{sensor-syncronsisation-interface} is set up. The
last step in the setup is to set up communication with the host or
connected \gls{pc}. All microcontroller platforms used here have a
\gls{usb} slave port. The used usb descriptor is
\passthrough{\lstinline!Serial over (+usb)!}-((+usb)\gls{cdc}). This is
used to emulate a virtual RS232 communication port using a \gls{usb}
port on a \gls{pc} and usually no additional driver is needed on modern
systems.

Now that the setup process is complete, the system switches to an
infinite loop, which processes several possible actions. One task is, to
react to user commands which can be sent to the system by the user via
the integrated \gls{cli}. All sensors are read out via a timer interval
set in the setup procedure and their values are stored in a ringbuffer.
Ring buffers, offer efficient data management in limited memory. Its
cyclic structure enables continuous overwriting of older data, saves
memory space and facilitates seamless processing of real-time data.

Ring buffers are well suited for applications with variable data rates
and minimise the need for complex memory management. The buffer can be
read out by command and the result of the measurement is sent to the
host. Each sensor measurement result is transmitted from the buffer to
the host together with a time stamp and a sequential number. This
ensures that in a multi-sensor setup with several sensors. The
measurements are synchronized\ref{sensor-syncronsisation-interface} in
time and are not out of sequence or drift.

@@

\begin{itemize}
\tightlist
\item
  Use with \gls{mrp}-library\ref{software-readout-framework}
\item
  Stand-alone mode via sending commands using built-in \gls{cli}
\end{itemize}

The \gls{cli} mode is a simple text-based interface with which it is
possible to read out current measured values, obtain debug information
and set operating parameters. This allows to quickly determine whether
the hardware is working properly after installation. The \gls{cli}
behaves like terminal programmes, displaying a detailed command
reference\ref{Sensors_(+cli).png} to the user after connecting. The
current measured value can be output using the
\passthrough{\lstinline!readout!}
command\ref{Query_sensors_b_value_using_(+cli).png}.

@@

The other option is to use the \gls{mrp}-library. The serial interface
is also used here. However, after a connection attempt by the
\passthrough{\lstinline!MRPHal!} module\ref{hal} of the
\gls{mrp}-library, the system switches to binary mode, which is
initiated using the \passthrough{\lstinline!sbm!} command. The same
commands are available as for \gls{cli}-based communication.

@@

\begin{longtable}[]{@{}llllll@{}}
\caption{Measured sensor readout to processing using host software
\label{Measured_sensor_readout_to_processing_using_host_software.csv}}\tabularnewline
\toprule
\begin{minipage}[b]{0.02\columnwidth}\raggedright
Run\strut
\end{minipage} & \begin{minipage}[b]{0.04\columnwidth}\raggedright
Points\strut
\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\raggedright
Host software runtime {[}ms{]}\strut
\end{minipage} & \begin{minipage}[b]{0.32\columnwidth}\raggedright
Average Sensor communication time per reading {[}ms{]}\strut
\end{minipage} & \begin{minipage}[b]{0.20\columnwidth}\raggedright
Communication jitter time {[}ms{]}\strut
\end{minipage} & \begin{minipage}[b]{0.09\columnwidth}\raggedright
Comment\strut
\end{minipage}\tabularnewline
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{0.02\columnwidth}\raggedright
Run\strut
\end{minipage} & \begin{minipage}[b]{0.04\columnwidth}\raggedright
Points\strut
\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\raggedright
Host software runtime {[}ms{]}\strut
\end{minipage} & \begin{minipage}[b]{0.32\columnwidth}\raggedright
Average Sensor communication time per reading {[}ms{]}\strut
\end{minipage} & \begin{minipage}[b]{0.20\columnwidth}\raggedright
Communication jitter time {[}ms{]}\strut
\end{minipage} & \begin{minipage}[b]{0.09\columnwidth}\raggedright
Comment\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.02\columnwidth}\raggedright
1\strut
\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedright
1\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright
9453\strut
\end{minipage} & \begin{minipage}[t]{0.32\columnwidth}\raggedright
1.44\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\raggedright
0\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.02\columnwidth}\raggedright
2\strut
\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedright
1\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright
9864\strut
\end{minipage} & \begin{minipage}[t]{0.32\columnwidth}\raggedright
1.5\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\raggedright
0\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.02\columnwidth}\raggedright
3\strut
\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedright
10\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright
12984\strut
\end{minipage} & \begin{minipage}[t]{0.32\columnwidth}\raggedright
1.22\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\raggedright
0.9\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.02\columnwidth}\raggedright
4\strut
\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedright
10\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright
12673\strut
\end{minipage} & \begin{minipage}[t]{0.32\columnwidth}\raggedright
1.13\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\raggedright
1.1\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.02\columnwidth}\raggedright
5\strut
\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedright
10\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedright
43264\strut
\end{minipage} & \begin{minipage}[t]{0.32\columnwidth}\raggedright
2.19\strut
\end{minipage} & \begin{minipage}[t]{0.20\columnwidth}\raggedright
8.2\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedright
96\% system load\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

@@

\begin{itemize}
\tightlist
\item
  \passthrough{\lstinline!Raspberry Pi OS Lite!} -
  \passthrough{\lstinline!debian bookworm x64!},
\item
  \gls{mrp}\ref{software-readout-framework}-library - version
  \passthrough{\lstinline!1.4.1!}
\item
  \passthrough{\lstinline!Unified Sensor Firmware!}\ref{unified-sensor}
  - version \passthrough{\lstinline!1.0.1!}
\end{itemize}

@@

The other issue is sending the trigger signal from the readout
software\ref{software-readout-framework}. Here too, unpredictable
latencies can occur, depending on which other tasks are also executed on
this port.

In order to enable the most stable possible synchronisation between
several sensors, an option has already been created to establish an
electrical connection between sensors. This is used together with the
firmware to synchronise the readout intervals. The
schematic\ref{Multi_sensor_synchronisation_wiring_example.png} shows how
several sensors must be wired together in order to implement this form
of synchronisation.

@@

Two functional sensor
platforms\ref{Build_sensors_with_different_capabilities.csv} were built
in order to create a solid test platform for later tests and for the
development of the \gls{mrp} library with the previously developed
sensor concepts.

@@

These cover all the required functions described in the
Usecases\ref{usecases}. The most important difference, apart from the
sensor used, is the \passthrough{\lstinline!scan mode!}. In this
context, this describes whether the sensor can measure a
\passthrough{\lstinline!static!} fixed point on the magnet or if the
sensor can move \passthrough{\lstinline!dynamically!} around the magnet
using a controllable manipulator.

@@

The \passthrough{\lstinline!1D!}
sensor\ref{1D_sensor_construction_with_universal_magnet_mount.png} is
the simplest possible sensor that is compatible with the
\passthrough{\lstinline!Unified Sensor Firmware!}\ref{firmware} platform
and also with the \gls{mrp}-library\ref{software-readout-framework}.

The electrical level here is based on a
\passthrough{\lstinline!Raspberry-Pi Pico!} together with the
\passthrough{\lstinline!MMC5603NJ!} magnetic sensor. The mechanical
setup consists of four 3D-printed components, which are fixed together
with nylon screws to minimise possible influences on the measurement.

>>>>

\hypertarget{shimming-procedure}{%
\subsection{Shimming Procedure}\label{shimming-procedure}}

@@

\ref{usecases}

@@

\begin{itemize}
\item
  Chapter \ref{state-of-the-art}. \textbf{State of the Art} describes
  the current state of technology and forms the basis for further
  development. Existing measurement methods and findings are shown here
  in order to define the context for the current project.
\item
  Chapter \ref{usecases}. \textbf{Usecases} are the application cases in
  which the project is to be used. They illustrate the practical
  scenarios and define how the product can be used in the real world.
  This is crucial for understanding the needs of the target group and
  designing the end result accordingly.
\item
  Chapter \ref{unified-sensor}. \textbf{Unified Sensor} refers to the
  integration of different sensors into a standardised solution. This
  enables simple data acquisition and serves as a basic hardware system
  on which the subsequent data processing library can be applied.
\item
  Chapter \ref{software-readout-framework}. \textbf{Software Readout
  Framework} describes the implementation of the data readout framework.
  This includes an explanation of the various modules and specific
  application examples.
\item
  Chapter \ref{usability-improvements}. \textbf{Usability Improvements}
  refers to additional activities to improve user-friendliness. This
  includes the optimisation of interfaces, interactions and processes to
  ensure intuitive and efficient use of the product. This also includes
  the documentation of code and the distribution of the source code as a
  package to users.
\item
  Chapter \ref{evaluation}. \textbf{Evaluation} comprises the systematic
  review and assessment of the overall system. This includes the
  demonstration of the implemented capabilities of the overall system
  against the previously defined usecases.
\end{itemize}

\hypertarget{state-of-the-art}{%
\chapter{State of the Art}\label{state-of-the-art}}

\hypertarget{opensource-projects}{%
\section{Opensource Projects}\label{opensource-projects}}

\hypertarget{conceptual-design}{%
\section{Conceptual Design}\label{conceptual-design}}

@@

accomplished

@@

The following section defines some usecases that the future project
should be able to cover. These illustrate practical situations to
understand the functionality and added value of the implemented solution
for the user. These were defined in the course of project planning and
provide an overview of how the user interacts with the project and what
functionalities can be expected. In the later accomplished evaluation
process \ref{evaluation}, the defined usecases are also used as a
reference to demonstrate the implemented capabilities of the solution.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Ready to use hardware sensor designs}
\end{enumerate}

A universal, easy-to-integrate Hall sensor design allows users to
evaluate the framework quickly and cost-effectively. The pre-built
hardware sensors provide an optimal solution for research, reducing
development time and achieving repeatable measurement results. Once
successfully evaluated, the Hall sensor design should be easily
adaptable to other sensors without the need for major firmware changes.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  \textbf{Taking automatic measurements from sensors}
\end{enumerate}

The purpose of the framework is to enable the automated acquisition of
measurement data from various connected hadware sensors. The user should
be able to configure various measurement series, which should then be
carried out by the framework without further user interaction.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{2}
\tightlist
\item
  \textbf{Open storage formats for data export}
\end{enumerate}

The use of open storage formats for the export of data enables an
interoperable data exchange environment. The implementation of
standardized formats improves the portability and long-term availability
of data. This encourages the exchange and further processing of
measurement data in other software tools.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{3}
\tightlist
\item
  \textbf{Ready to integrate data analysis functions}
\end{enumerate}

Once it is possible to record and store measured values, it should be
possible for the user to analyse and visualize this data using various
algorithms. The focus here should be on extending the framework with
user-created algorithms.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{4}
\tightlist
\item
  \textbf{User programmable data processing pipelines}
\end{enumerate}

User-programmable data processing pipelines enable the flexible design
of data processing sequences as pipeline by users. The framework should
enable users to create their own pipelines with the previously defined
data analysis functions.

@@

A defined main objective of this project is the development of a
cost-effective magnetic field sensors interface that is also universally
expandable. The focus is on mapping different sensors and being
compatible with different magnet types and shapes. This ensures broad
applicability in different scenarios. Another goal is reproducibility to
ensure consistent results. Easy communication with standard PC hardware
maximises the user-friendliness of the interface. The flexibility to
support different sensors and magnets makes the system versatile and
opens the possibility for use in different applications. A low-cost
magnetic field sensors interface will therefore not only be economically
attractive, but also facilitate the integration of magnetic field
sensors in different contexts. In addition, the low-cost sensor
interface will serve as a development platform for the data evaluation
\gls{mrp} library and provide real measurement data from magnets. In
addition, the interface firmware creates a basis for the development of
a data protocol for exchanging measured values. This makes it easy to
integrate your own measuring devices into the \gls{mrp} ecosystem at a
later date. This is only possible with a minimal functional hardware and
firmware setup and was developed for this purpose first.

\hypertarget{sensor-selection}{%
\section{Sensor Selection}\label{sensor-selection}}

@@

The use of analog sensors was purposefully avoided, though they are
suitable for more precise measurements and extended measuring ranges.
They were excluded because they require more carefully designed circuits
and more complicated energy management. In the context of the desired
goal of developing a cost-efficient and universally expandable Hall
sensor interface, the decision in favour of digital sensors seems
appropriate.

Focussing on the digital \gls{i2c} interface not only facilitates
implementation, but also contributes to overall cost efficiency. At the
same time, the integration of temperature sensors enables precise
measurements under varying environmental conditions. This strategic
choice forms the basis for a flexible, universally applicable Hall
sensor interface that can be seamlessly integrated into various existing
systems.

\begin{longtable}[]{@{}lllll@{}}
\caption{Implemented digital magnetic field sensors
\label{Implemented_digital_magnetic_field_sensors.csv}}\tabularnewline
\toprule
& TLV493D-A1B6 & HMC5883L & MMC5603NJ & AS5510\tabularnewline
\midrule
\endfirsthead
\toprule
& TLV493D-A1B6 & HMC5883L & MMC5603NJ & AS5510\tabularnewline
\midrule
\endhead
Readout-Axis & 3D & 3D & 3D & 1D\tabularnewline
Temperature-Sensor & yes & no & yes & no\tabularnewline
Resolution {[}uT{]} & 98 & 0.2 & 0.007 & 48\tabularnewline
Range {[}mT{]} & ±130.0 & ±0.8 & ±3 & ±50\tabularnewline
Interface & \gls{i2c} & \gls{i2c} & \gls{i2c} & \gls{i2c}\tabularnewline
\bottomrule
\end{longtable}

The table \ref{Implemented_digital_magnetic_field_sensors.csv} shows a
selection of sensors for which hardware and software support has been
implemented. The resolution of the selected sensors covers the expected
range of values required by the various magnets to be tested.

@@

The mechanical design of a sensor would be kept as simple as possible so
that it can be replicated as easily as possible. The focus was on
providing a stable foundation for the sensor \gls{ic} and an
exchangeable holder for different magnets.

The following figure
\ref{1D_sensor_mechanical_3D_printed_structure.png}, shows a sectional
view of the \gls{cad} drawing of the 1D-Single sensor
\ref{d-single-sensor}.

All parts were produced using the 3D printing additive manufacturing
processes. The sensor circuit board was glued underneath the magnet
holder. This is interchangeable, so different distances between sensor
and magnet can be realised.

The exchangeable magnetic holder (shown in green) can be adapted to
different magnets. It can be produced quickly due to the small amount of
parts used. The two recesses lock the magnet holder with the inserted
magnet over the sensor. The specified tolerances allow the magnet to be
inserted into the holder with repeat accuracy and without backlash. This
is important if several magnets have to be measured, where the
positioning over the sensor must always be the same.

@@

The electronics consist of the magnetic field sensor and the electrical
interface to connect it to a \gls{pc} in the form of a microcontroller.

The focus here was on utilising existing microcontroller development and
evaluation boards, which already integrate all the components required
for basic operation. This not only enabled a time-saving implementation,
but also ensured a cost-efficient realisation.

All the necessary components and their circuitry were then recorded on a
\gls{pcb} \ref{1D_sensor_schematic_and_circuit_board.png} and
subsequently manufactured. In addition, footprints were provided for
various sensor \gls{ic} packages. By placing mounting holes on the
\gls{pcb}, it is possible to attach various mechanical mounts ontop of
the sensor \gls{ic}s.

Special attention was paid to the provision of an accessible
SYNC-\gls{gpio} connector. This enables subsequent multi-sensor
synchronization and also offers options for later extensions. This
functionality opens up the possibility of synchronising data from
different sensors to achieve precise and coherent measurement results.
Overall, this integrated approach represents an effective solution for
the flexible evaluation of sensors and helps to optimise the development
process.

@@

The microcontroller firmware is software that is executed on a
microcontroller in an embedded system. It controls the hardware and
enables the execution of predefined functions. The firmware is used to
process input data, control output devices and performs specific tasks
according to the program code. It handles communication with sensors,
actuators and other peripheral devices, processing data and making
decisions. Firmware is critical to the functioning of devices.

@@

The firmware is responsible for detecting the possible connected sensors
\ref{Implemented_digital_magnetic_field_sensors.csv} and query
measurements. This measured data can be forwarded to a host \gls{pc} via
a user interface and can then be further processed there.

An important component is that as many common sensors as possible can be
easily connected without having to adapt the firmware. This modularity
was implemented using abstract class design. These are initiated
according to the sensors found at startup. If new hardware is to be
integrated, only the required functions \ref{lst:CustomSensorClass} need
to be implemented.

@@

The flow chart
\ref{Unified_sensor_firmware_simplified_program_strucutre.png} shows the
start process and the subsequent main loop for processing the user
commands and sensor results. When the microcontroller is started, the
software checks whether known sensors are connected to \gls{i2c} or
\gls{uart} interfaces. If any are found (using a dedicated \gls{lut}
with sensor address translation information), the appropriate class
instances are created and these can later be used to read out
measurement results.

The next initialisation system is dedicated for multi-sensor
synchronisation \ref{sensor-syncronsisation-interface}. The last step in
the setup is to configure communication with the host or connected
\gls{pc}. All implemented microcontroller platforms used (Raspberry Pi
Pico, STM32F4) have a \gls{usb} slave port. The used usb descriptor is a
\gls{usb} \gls{cdc}. This is used to emulate a virtual RS232
communication port using a \gls{usb} port on a \gls{pc} and usually no
additional driver is needed on modern host systems.

After the setup process is complete, the system switches to an infinite
loop, which processes several possible actions. One task is, to react to
user commands which can be sent to the system by the user via the
integrated \gls{cli}. All sensors are read out via a timer interval set
in the setup procedure and their values are stored in a ringbuffer. Ring
buffer offers efficient data management in limited memory. Its cyclic
structure enables continuous overwriting of older data, saves memory
space and facilitates seamless processing of real-time data.

Ring buffers are well suited for applications with variable data rates
and minimise the need for complex memory management. The buffer can be
read out by command and the result of the measurement is sent to the
host. Each sensor measurement result is transmitted from the buffer to
the host together with a time stamp and a sequential number. This
ensures that in a multi-sensor setup with several sensors. The
measurements are synchronized \ref{sensor-syncronsisation-interface} in
time and are not out of sequence or drift.

@@

\begin{itemize}
\tightlist
\item
  Use with \gls{mrp} \ref{software-readout-framework}-libaray
\item
  Stand-alone mode via sending commands using built-in \gls{cli}
\end{itemize}

The \gls{cli} mode is a simple text-based interface with which it is
possible to read out current measured values, obtain debug information
and set operating parameters. This allows to quickly determine whether
the hardware is working properly after installation. The \gls{cli}
behaves like terminal programmes, displaying a detailed command
reference \ref{Sensors_(+cli).png} to the user after connecting. The
current measured value can be output using the
\passthrough{\lstinline!readout!}
command\ref{Query_sensors_b_value_using_(+cli).png}.

@@

The other option is to use the \gls{mrp}
\ref{software-readout-framework}-library. The serial interface is also
used here. However, after a connection attempt by the
\passthrough{\lstinline!MRPHal!}\ref{hal} module of the \gls{mrp}
\ref{software-readout-framework}-library, the system switches to binary
mode, which is initiated using the \passthrough{\lstinline!sbm!}
command. The same commands are available as for \gls{cli}-based
communication.

@@

\begin{longtable}[]{@{}lllll@{}}
\caption{Measured sensor readout to processing using host software
\label{Measured_sensor_readout_to_processing_using_host_software.csv}}\tabularnewline
\toprule
\begin{minipage}[b]{0.08\columnwidth}\raggedright
Data-Points\strut
\end{minipage} & \begin{minipage}[b]{0.11\columnwidth}\raggedright
Run runtime {[}ms{]}\strut
\end{minipage} & \begin{minipage}[b]{0.35\columnwidth}\raggedright
Average Sensor communication time per reading {[}ms{]}\strut
\end{minipage} & \begin{minipage}[b]{0.22\columnwidth}\raggedright
Communication jitter time {[}ms{]}\strut
\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright
Comment\strut
\end{minipage}\tabularnewline
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{0.08\columnwidth}\raggedright
Data-Points\strut
\end{minipage} & \begin{minipage}[b]{0.11\columnwidth}\raggedright
Run runtime {[}ms{]}\strut
\end{minipage} & \begin{minipage}[b]{0.35\columnwidth}\raggedright
Average Sensor communication time per reading {[}ms{]}\strut
\end{minipage} & \begin{minipage}[b]{0.22\columnwidth}\raggedright
Communication jitter time {[}ms{]}\strut
\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright
Comment\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.08\columnwidth}\raggedright
1\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
9453\strut
\end{minipage} & \begin{minipage}[t]{0.35\columnwidth}\raggedright
1.44\strut
\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright
0\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.08\columnwidth}\raggedright
1\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
9864\strut
\end{minipage} & \begin{minipage}[t]{0.35\columnwidth}\raggedright
1.5\strut
\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright
0\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.08\columnwidth}\raggedright
10\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
12984\strut
\end{minipage} & \begin{minipage}[t]{0.35\columnwidth}\raggedright
1.22\strut
\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright
0.9\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.08\columnwidth}\raggedright
10\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
12673\strut
\end{minipage} & \begin{minipage}[t]{0.35\columnwidth}\raggedright
1.13\strut
\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright
1.1\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.08\columnwidth}\raggedright
10\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright
43264\strut
\end{minipage} & \begin{minipage}[t]{0.35\columnwidth}\raggedright
2.19\strut
\end{minipage} & \begin{minipage}[t]{0.22\columnwidth}\raggedright
8.2\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright
96\% system load\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

@@

\begin{itemize}
\tightlist
\item
  Raspberry Pi OS Lite - \passthrough{\lstinline!debian bookworm x64!},
\item
  \gls{mrp} \ref{software-readout-framework}-library - version
  \passthrough{\lstinline!1.4.1!}
\item
  Unified Sensor \ref{unified-sensor}-firmware - version
  \passthrough{\lstinline!1.0.1!}
\end{itemize}

@@

The other issue is sending the trigger signal from the readout software
\ref{software-readout-framework}. Here too, unpredictable latencies can
occur, depending on which other tasks are also executed on this port.

In order to enable the most stable possible synchronisation between
several sensors, an option has already been created to establish an
electrical connection between sensors. This is used together with the
firmware to synchronise the readout intervals. The schematic
\ref{Multi_sensor_synchronisation_wiring_example.png} shows how several
sensors must be wired together in order to implement this form of
synchronisation.

@@

Two functional sensor platforms
\ref{Build_sensors_with_different_capabilities.csv} were built in order
to create a solid test platform for later tests and for the development
of the \gls{mrp} \ref{software-readout-framework}-library with the
previously developed sensor concepts.

@@

These cover all the required functions described in the usecases
\ref{usecases}. The most important difference, apart from the sensor
used, is the \passthrough{\lstinline!scan mode!}. In this context, this
describes whether the sensor can measure a
\passthrough{\lstinline!static!} fixed point on the magnet or if the
sensor can move \passthrough{\lstinline!dynamically!} around the magnet
using a controllable manipulator.

@@

The 1D sensor
\ref{1D_sensor_construction_with_universal_magnet_mount.png} is the
simplest possible sensor that is compatible with the Unified Sensor
firmware \ref{firmware} platform.

The electrical level here is based on a
\passthrough{\lstinline!Raspberry-Pi Pico!} together with the
\passthrough{\lstinline!MMC5603NJ!} magnetic sensor. The mechanical
setup consists of four 3D printed components, which are fixed together
with nylon screws to minimise possible influences on the measurement.
